---
title: "Sample Code in R"
author: "Md Mahi Uddin"
date: "October 5th, 2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,echo=F,fig_height=10,fig_width=7,cache = F)
```

I have solved following problems from the text book R Handbook and stated.

1. Question 1.1, pg. 23 in Handbook \textit{this question will require you to make some assumptions. List your assumptions and how you interpreted the question.}

```{r echo = T}
require(HSAUR3)
require(plyr)
require(dplyr)
require(MASS)
require(stringr)
require(plotrix)
require(ggplot2)
require(ISLR)
library(HSAUR3)
library(plyr)
library(MASS)
library(stringr)
library(plotrix)
library(ggplot2)
library(ISLR)
library(HSAUR2)
library(dplyr)
library(knitr)
data("Forbes2000", package = "HSAUR2")
USA_Companies<- subset(Forbes2000, country == "United States")
median(USA_Companies$profits, na.rm = TRUE)
UK_Companies<- subset(Forbes2000, country == "United Kingdom")
median(UK_Companies$profits, na.rm = TRUE)
France_Companies<- subset(Forbes2000, country == "France")
median(France_Companies$profits, na.rm = TRUE)
Germany_Companies<- subset(Forbes2000, country == "Germany")
median(Germany_Companies$profits, na.rm = TRUE)

#ampther method

mprofits <- tapply(Forbes2000$profits, Forbes2000$country, median, na.rm = TRUE)
median_Countries_profit <- mprofits[c("United Kingdom", "France", "Germany", "United States")]
median_Countries_profit


```



2. Question 1.2, pg. 23 in Handbook

```{r echo = T}
negative_profit_Germany<- subset(Forbes2000, country == "Germany" & profits < 0,select = c(name, profits))
negative_profit_Germany

```


3. Question 1.3, pg. 23 in Handbook

```{r echo = T}

Most_business_cat_Bermuda<- subset(Forbes2000, country == "Bermuda", select= c(name, category, country))
summary(Most_business_cat_Bermuda)
rev(sort(Most_business_cat_Bermuda$category))


```



4. Question 1.4, pg. 23 in Handbook


```{r echo = T}

head(sort(Forbes2000$profits, decreasing = TRUE))

fifty_ordr_profit<- Forbes2000[1:50, ]

plot(sales ~ assets, data =fifty_ordr_profit,xlim = c(0,1200), main = "Sales vs Assets", ylim=c(0,300))

labs=abbreviate(fifty_ordr_profit$country, minlength=2)
text(fifty_ordr_profit$assets, fifty_ordr_profit$sales,labs, col=2, font=3, pos = 3)

```




5. Question 1.5, pg. 23 in Handbook


```{r echo = T}
#Find the average value of sales for the companies in each country in the Forbes data set
#package:base

mean_sale<- aggregate(sales~country, mean, data = Forbes2000)
mean_sale
#msale <- tapply(Forbes2000$sales, Forbes2000$country, mean, na.rm = TRUE)
#Ave_sales1 <- data.frame(Avg=msale)
#Ave_sales1

#and ???nd the number of companies in each country with pro???ts above 5 billion US dollars

No_Five_bill=subset(Forbes2000, profits>5, select = c(country,profits))
No_Five_bill


```




6. Question 2.1, pg. 39 in Handbook (see Chapter 6 of R Graphcis Cookbook for GGPlot)

```{r echo = T}
data("household")
library(HSAUR3)
attach(household)
HH_dat<- household
Tot_Expen <- apply(HH_dat[,1:4],1,sum)
Cbind_data<- cbind(HH_dat, Tot_Expen)
pairs(Cbind_data[,-5],col=rainbow(5)[Cbind_data$gender],cex=1.5,lty=4)


#total expenditure
ggplot(Cbind_data, aes(x=gender, y=Tot_Expen, fill = gender)) + geom_boxplot(width=.2)+ggtitle("Boxplot of gender and Total Expenditure") + labs(x="Gender",y="Total Expenditure") 


#housing
ggplot(Cbind_data, aes(x=gender, y=housing, fill = gender)) + geom_boxplot(width=.2)+ggtitle("Boxplot of gender and Expenditure on Housing") + labs(x="Gender",y="Housing") 



#food
ggplot(Cbind_data, aes(x=gender, y=food, fill = gender)) + geom_boxplot(width=.2)+ggtitle("Boxplot of gender and Expenditure on food") + labs(x="Gender",y="food") 


#goods
ggplot(Cbind_data, aes(x=gender, y=goods, fill = gender)) + geom_boxplot(width=.2)+ggtitle("Boxplot of gender and Expenditure on goods") + labs(x="Gender",y="goods") 


#service
ggplot(Cbind_data, aes(x=gender, y=service, fill = gender)) + geom_boxplot(width=.2)+ggtitle("Boxplot of gender and Expenditure on service") + labs(x="Gender",y="service") 


```


7. Question 2.3, pg. 41 in Handbook (see Chapter 6 of R Graphcis Cookbook for GGPlot)


```{r echo = T}
data("suicides2", package = "HSAUR3")
head(suicides2)
plot(ecdf(suicides2$A25.34))
rug("suicides2$A25.34")

truehist(suicides2$A35.44)
rug(suicides2$A35.44)

ggplot(suicides2, aes(x=suicides2$A45.54)) + geom_density(fill="blue", alpha=0.2)
ggplot(suicides2, aes(x=suicides2$A55.64)) + geom_density(fill="green", alpha=0.2)
ggplot(suicides2, aes(x=suicides2$A65.74)) + geom_density(fill="yellow", alpha=0.2)


```


8. Using a single R expression, calculate the median absolute deviation, $1.4826\cdot median|x-\mu|$, where $\mu$ is the sample median. Use the dataset \textbf{chickwts}. Use the R function mad() to verify your answer.

```{r echo = T}
attach(chickwts)
chickwts
weight_dataset <- chickwts$weight
weight_dataset

#calculating the median of weight_dataset
median_weight <- median(weight_dataset)
median_weight
head(median_weight)
# calculate the median absolute deviation (MAD)
MAD <- abs(weight_dataset - median_weight)
MAD
calculated_output <- 1.4826 * median(MAD)
calculated_output

# Use the R function mad() to verify your answer
mad(weight_dataset, median_weight)
#the answer of both the methods are similar

```


9. Using the data matrix \textbf{state.x77}, obtain side-by-side boxplots of the per capita income variable for the nine different divisions defined by the variable \textit{state.division}. Comment on the plot.

```{r echo = T}
head(state.x77)
#state.x77
# drawing the boxplot per capital income for nine different divisions
head(state.division)
#state.division
state_div_cbind <- cbind(state.x77, state.division)
#data_set_state


boxplot(Income ~ state.division, data = state_div_cbind, col = c("red","blue","yellow","green", "sienna", "royalblue", "aquamarine","cyan", "deepskyblue"), main = "Per capita income by State Division")


boxplot(Income ~ state.division, data = state_div_cbind, names=c("New Eng", "Middle Atl.", "South Atl.", "E.S. Central", "W.S. Central","E.N. Central", "W.N. Central", "Mountain", "Pacific"), las= 2,cex.axis = 0.7, col = c("2","3","4","5", "6", "7", "10","11", "12"), main = "Per capita income by State Division")
## We can see that besides Middle Atl. 
# Most have similar medians. East South Central and West North Central are skewed to the left.

```


10. Using the data matrix \textbf{state.x77}, find the state with the minimum per capita income in the New England region as defined by the factor \textit{state.division}. Use the vector \textit{state.name} to get the state name.

```{r echo = T}
state_divisions_cbine <- data.frame(state.x77, state.division)

New_England_Div <- subset(state_divisions_cbine, state.division == "New England")

order(New_England_Div$Income)
New_England_ordr_income <- New_England_Div[order(New_England_Div$Income, decreasing = T)]
New_England_ordr_income

```


11. Use subscripting operations on the dataset \textbf{Cars93} to find the vehicles with highway mileage of less than 25 miles per gallon (variable \textit{MPG.highway}) and weight (variable \textit{Weight}) over 3500lbs. Print the model name, the price range (low, high), highway mileage, and the weight of the cars that satisfy these conditions.

```{r echo = T}
data(Cars93, package = "MASS")

selected_info_cars <- subset(Cars93, MPG.highway < 25 & Weight > 3500, c(Model,MPG.highway, Weight, 
   Min.Price, Price, Max.Price))
selected_info_cars
```





12. Form a matrix object named \textbf{mycars} from the variables \textit{Min.Price, Max.Price, MPG.city, MPG.highway, EngineSize, Length, Weight} from the \textbf{Cars93} dataframe from the \textbf{MASS} package. Use it to create a list object named \textit{cars.stats} containing named components as follows:

    a) A vector of means, named \textit{Cars.Means}
    
```{r echo = T}

mycars <- as.matrix(Cars93[, c("Min.Price", "Max.Price", "MPG.city", "MPG.highway","EngineSize", "Length","Weight")])
#mycars
#head(mycars)

Cars.Means <- colMeans(mycars, na.rm = FALSE, dims = 1)
Cars.Means

# Find: vector of means, named Cars.Means
#Cars.Means <- apply(mycars, 2, mean,na.rm = FALSE)
#Cars.sd <- apply(mycars, 2, sd)
#head(Cars.Means)
#head(Cars.sd)
#std.error(mycars,na.rm)

```
    
    
    b) A vector of standard errors of the means, named \textit{Cars.Std.Errors}
    
```{r echo = T}
#std.error(mycars, na.rm)


#Use describe function from psych package for standard error
library(psych)
find_se <- describe(mycars, type=2)
Cars.Std.Error <- find_se$se 
#extract_std_error
#Cars.Std.Error <- extract_std_error
#Cars.Std.Error

names(Cars.Std.Error)<- c("Min.Price", "Max.Price", "MPG.city", "MPG.highway","EngineSize", "Length","Weight")
#head(Cars.sd)
Cars.Std.Error

```
    
    c) A matrix with two rows containing lower and upper limits of 99% Confidence Intervals for the means, named \textit{Cars.CI.99}
    
```{r echo = T}

confidence_level <- function(x)t.test(x, conf.level = 0.9)$conf.int
Cars.CI.99 <- as.matrix(apply(mycars, 2, confidence_level))
Cars.CI.99
row.names(Cars.CI.99) <- c("Lower Limit", "Upper Limit")
Cars.Stats <- list(Cars.Means,Cars.Std.Error,Cars.CI.99)
names(Cars.Stats) <- c("Cars.Means","Cars.Std.Error","Cars.CI.99")
Cars.Stats

```
    
    
13. Use the \texttt{apply()} function on the three-dimensional array \textbf{iris3} to compute:

    a) Sample means of the variables \textit{Sepal Length, Sepal Width, Petal Length, Petal Width}, for each of the three species \textit{Setosa, Versicolor, Virginica}
    
```{r echo = T}


dataset_iris3 <- iris3
head(dataset_iris3)
#iris3


# sample means of the variables
#Sepal Length, Sepal Width, Petal Length, and Petal Width for each of the three species
#Setosa, Versicolor and Virginica
specific_dataset <- apply(dataset_iris3, c(2,3), mean)
head(specific_dataset)

```
    
    
    b) Sample means of the variables \textit{Sepal Length, Sepal Width, Petal Width} for the entire data set.
    
```{r echo = T}


# variables Sepal Length, Sepal Width, Petal Length, and Petal Width, 
# for the entire data set
entire_dataset <- apply(dataset_iris3, c(1,2), mean)
entire_dataset
head(entire_dataset)

```
    
    
14. Use the data matrix \textbf{state.x77} and the \texttt{tapply()} function to obtain:

    a) The mean per capita income of the states in each of the four regions defined by the factor \textit{state.region}
```{r echo = T}



data(state)
head(state.x77)
dataset_state_x77 <- state.x77

# Problem 9a
# the mean per capital income of the states in each of the 
# four regions defined by the factor state.region
mean_per_cap_income <-tapply(dataset_state_x77[,"Income"], state.region, mean)
mean_per_cap_income

```
    
    
    b) The maximum illiteracy rates for states in each of the nine divisions defined by the factor \textit{state.division}
    
```{r echo = T}

max_illiteracy_rate <- tapply(dataset_state_x77[,"Illiteracy"], state.division, max)
max_illiteracy_rate

```
    
    
    c) The number of states in each region
    
```{r echo = T}

new_dataset_state <- cbind(state.x77, state.region)
#head(new_dataset_state)
num_state_each_region <- tapply(new_dataset_state[,"state.region"], state.region, sum)
num_state_each_region

```
    
    
    d)  The median high school graduation rates for groups of states defined by combinations of the factors \textit{state.region} and \textit{state.size}
    
        ```{r, eval=F}
        state.size = cut(state.x77[,"Population"],breaks=c(0,2000,10000,Inf),
        +labels=c("Small","Medium","Large"))
        
        ```
```{r echo = T}

state.size <- cut(state.x77[,"Population"], breaks = c(0, 2000, 10000, Inf),  labels = c("Small", 
"Medium", "Large"))

#head(state.size)

data_frame_state_size <- cbind(new_dataset_state, state.size)
head(data_frame_state_size)
Data_HS_Grad_1 <- tapply(data_frame_state_size[, "HS Grad"], list(state.region, state.size), median)
Data_HS_Grad_1

#Data_HS_Grad_2 <- tapply(new_dataset_state[, "HS Grad"],new_dataset_state[, "state.region"], median)
#Data_HS_Grad_2

```
    

15. Using the dataframe \textbf{mtcars}, produce a scatter plot matrix of the variables \textit{mpg, disp, hp, drat, qsec}. Use different colors to identify cars belonging to each of the categories defined by the \textit{carsize} variable in different colors.

    ```{r echo = T, eval=F}
carsize = cut(mtcars[,"wt"], breaks=c(0, 2.5, 3.5, 5.5), labels = c("Compact","Midsize","Large"))
    
    ```
    
```{r echo = T}


data(mtcars)
head(mtcars)

carsize <- cut(mtcars[, "wt"], breaks=c(0, 2.5, 3.5, 5.5), labels=c("Compact", "Midsize", "Large"))
#carsize
mtcars1 <- cbind(mtcars, carsize)
#data(mtcars)
head(mtcars1)
pairs(~mpg+disp+hp+drat+qsec,data=mtcars1, col=as.integer(mtcars1$carsize),
      main="Scatterplot of selected variables")

```

16. Use the function \texttt{aov()} to perform a one-way analysis of variance on the \textbf{chickwts} data with \textit{feed} as the treatment factor. Assign the result to an object named \textit{chick.aov} and use it to print an ANOVA table. Use this object to obtain side-by-side box plots of the residuals for each feed.

```{r echo = T}


# side-by-side box plots of the residuals for each feed
attach(chickwts)
head(chickwts)


chick.aov <- aov(weight ~ feed, data=chickwts)
summary(chick.aov)
print(chick.aov)
residuals <- resid(chick.aov)
boxplot(residuals ~ feed, data = cbind(chickwts,residuals),las= 2,cex.axis = 1, col = c("2","3","4","5", "6", "7"), main = "Residuals for each feed")

```


17. Write an R function named \texttt{ttest()} for conducting a one-sample t-test. Return a list object containing the two components: 

    - the t-statistic named T;
    
    - the two-sided p-value named P.
    
Use this function to test the hypothesis that the mean of the \textit{weight} variable (in the \textbf{chickwts} dataset) is equal to 240 against the two-sided alternative. \textit{For this problem, please show the code of function you created as well as show the output. }

```{r echo = T}

attach(chickwts)
#head(chickwts)
mu = 240
ttest <- function(x) (mean(x)-mu)/(sd(x)/sqrt(length(x)))
ttest(weight)
#mean(weight)
#sd(weight)

t.test(weight,alternative = c("two.sided"),mu=240) # mu is the mean weight

```


    
